```{python echo=FALSE}
import warnings
import pandas as pd
import numpy as np
import python.api_calls as api
import python.connectivity_assess as conn_table

warnings.filterwarnings('ignore')
```

```{r echo=FALSE}
#| echo: false
#| include: true
#| warning: false

library(reticulate)
library(flextable)


```

# Methods and Definitions

## Connectivity Modelling {-}
To prioritize culverts for remediation, connectivity status and structure ranks were produced using BCFishPass, an open-source freshwater connectivity modelling framework, which builds off provincial barrier mapping tools (Mount et al. 2011) and includes three sub-models: 

Accessibility model: Naturally accessible waterbodies are those that are considered likely accessible to focal species if no human-made barriers existed on the landscape. These were spatially delineated for each focal species using natural barriers (i.e., waterfalls, gradient barriers, or subsurface flows) that would naturally limit upstream movement. 

Habitat model: A subset of the naturally accessible waterbody layer was defined as key habitat, i.e., habitat likely to support spawning or rearing, rather than simply being used as a movement corridor. The habitat model identifies areas that have a higher potential to support key habitat based on stream channel gradient and discharge. 

Connectivity model: A layer of known or modelled structures was overlaid on the key habitat model output. Modelled stream crossings were created from stream network and road/railway Geographic Information System (GIS) data, by mapping where a road or railway crosses a stream. Dam data were obtained from the Canadian Aquatic Barriers Database. All mapped dams and modeled crossings (henceforth structures) downstream of key habitat were initially considered potential barriers. 

Some structures were classified as non-existent or passable prior to initiating field work and excluded from the model. Stream crossings on 6th order and larger streams were assumed to be bridges and classified as passable.  Remaining structures were checked on Google Earth to verify they were not bridges, cross-ditches, or non-existent crossings. Data from previous field assessments was obtained from the Provincial Stream Crossing Inventory System (PSCIS), and structures that had been assessed and scored as passable were classified as such. Assessed structures that were scored as barriers or potential barriers were considered barriers in the model, along with all dams lacking fishways and structures that have not yet been assessed. Models were run to produce initial outputs, and maps were created showing disconnected key habitat and associated structures. 

Local knowledge and data were incorporated into model development in two phases and overruled modeled outputs where the two differed. Existing reports on habitat, barriers, and fish distributions, when available, were incorporated into the initial model outputs. Maps of initial outputs were then shared with local knowledge holders, who identified model errors or omissions (e.g., passable structures, or known spawning habitats that were not identified by the model). Corrections to model data were made based on local knowledge, and models were re-run. 

Connectivity status was estimated by calculating the proportion of key habitat that is connected to the ocean (i.e., not fragmented by human-made barriers). Key habitat with no structures or only passable structures downstream was considered connected.  

Ranking: Structures, and sets of structures that combine to disconnect habitat, were ranked by the amount of key habitat upstream. A ranked structure list helps determine where structures are blocking large amounts of habitat, and where removal of an individual structure or a group of structures (a “set”) is more effective in opening up habitat, by ranking structures based on amount of habitat blocked. Top-ranked structures or sets of structures represented those with the most key habitat upstream and did not consider other factors such as habitat quality and cost, which ultimately drove restoration prioritization when planning field work and deciding priorities with partners.  

Habitat Accumulation Curves (HACs): HACs are used to visualize structure rankings in order of the relative amount of habitat they are blocking upstream. Structures are plotted in order of km’s blocked by structures. HACs illustrate restoration priorities by showing how much habitat becomes accessible as barriers are removed or improved. This method of visualizing cumulative habitat gains is useful for prioritizing restoration work and communicating results to partners. 


```{python echo=FALSE}
#| warning: false
#| echo: false

#| label: tbl-connectivity
#| tbl-cap: "Connectivity status assessment for thermal refuge (a), spawning (b), and rearing (c) habitat in the Lower Nicola River watershed. The Accessible Spawning Habitat and Accessible Rearing Habitat KEAs are evaluated by dividing the length of linear habitat (of each type) that is currently accessible to focal species by the total length of all linear habitat (of each type) in the watershed."
#| warning: false
#| echo: false

df = conn_table.make_assess_table("Anadromous salmonids", "Available off-channel Thermal Refuge", "Total Area (m2) of thermal refuge accessible", "?", " ?", "? ", api.connect)
df


```

**Comments:** No baseline data exists on the extent of overwintering habitat in the watershed. A priority action is included in the Operational Plan (strategy 2.3) to develop a habitat layer, and this will be used to inform this connectivity status assessment in the future.

```{python echo=FALSE}
#| warning: false
#| echo: false
df = conn_table.make_assess_table("Anadromous salmonids", "Available Spawning Habitat", "% of total linear spawning habitat accessible", 25, 50, 75, api.connect_spawn)
df

```



**Comments:** Indicator rating definitions are based on the consensus decisions of the planning team, including the decision not to define Fair. The current status is based on the connectivity model output, which is current as of December 2024.

```{python echo=FALSE}

df = conn_table.make_assess_table("Anadromous salmonids", "Available Rearing Habitat", "% of total linear rearing habitat accessible", 25, 50, 75, api.connect_rear)
df

```
**Comments:** Indicator rating definitions are based on the consensus decisions of the planning team, including the decision not to define Fair. The current status is based on the connectivity model output, which is current as of December 2024.


## Structure Ranking {-}
### Lateral Barriers (Including to thermal refugia) {.unnumbered}

There is a lack of comprehensive data and mapping of lateral barriers and potential thermal refugia in the watershed to currently support a strategic prioritization (see @tbl-S1). However, local knowledge was used to compile a list of candidate sites for field verification as a starting point to improve lateral and thermal connectivity. These sites were investigated in 2021 (@Mozin2021), but an atmospheric river event later that winter resulted in major modifications and rearrangement of the Coldwater River and mainstem Nicola River downstream of the Coldwater confluence, rendering the outcome of those investigations no longer relevant. Further assessment of off-channel thermal refugia was undertaken by Nooaitch Band in 2023, with results of those assessments expected to be released to the broader community working on conservation in the Lower Nicola Watershed in the near future.  

### Longitudinal Field Assessment Ranking Process {.unnumbered}

There is a lack of comprehensive data and mapping of lateral barriers and potential thermal refugia in the watershed to currently support a strategic prioritization (see @tbl-S1). However, local knowledge was used to compile a list of candidate sites for field verification as a starting point to improve lateral and thermal connectivity. These sites were investigated in 2021 (@Mozin2021), but an atmospheric river event later that winter resulted in major modifications and rearrangement of the Coldwater River and mainstem Nicola River downstream of the Coldwater confluence, rendering the outcome of those investigations no longer relevant. Further assessment of off-channel thermal refugia was undertaken by Nooaitch Band in 2023, with results of those assessments expected to be released to the broader community working on conservation in the Lower Nicola Watershed in the near future.  

```{python echo=FALSE}
#| label: tbl-table16
#| tbl-cap: "Rearing habitat connectivity gain requirements to meet WCRP goals in the Lower Nicola River watershed. The measures of currently accessible and total habitat values are derived from the intrinsic potential habitat model"
#| warning: false
#| echo: false

# Variables to API calls
connect = api.watershed_connectivity("ALL")[0]
total = round(api.watershed_connectivity("ALL")[1],2) 
access = round(api.watershed_connectivity("ALL")[2],2)
gain = round((total*0.96)-access,2)

# Data frame with table info
df = pd.DataFrame({"Habitat Type":["Spawning and Rearing"],
                   "Currently connected (km)":[str(access)],
                   "Total (km)": [str(total)],
                   "Current Connectivity Status":[str(connect)+"%"],
                   "Goal": ["90%"],
                   "Gain required (km)": [str(gain)]
                   })
                   
# Apply table styling
styled_df = df.style.hide().set_properties(**{'text-align': 'left'})
# header styling
table_styles = [{
       'selector': 'th',
       'props': [('background-color', '#008270'),('text-align', 'left')]
   }]
styled_df.set_table_styles(table_styles)

```

### Longitudinal Structure Prioritization Summary {.unnumbered}

There is a lack of comprehensive data and mapping of lateral barriers and potential thermal refugia in the watershed to currently support a strategic prioritization (see @tbl-S1). However, local knowledge was used to compile a list of candidate sites for field verification as a starting point to improve lateral and thermal connectivity. These sites were investigated in 2021 (@Mozin2021), but an atmospheric river event later that winter resulted in major modifications and rearrangement of the Coldwater River and mainstem Nicola River downstream of the Coldwater confluence, rendering the outcome of those investigations no longer relevant. Further assessment of off-channel thermal refugia was undertaken by Nooaitch Band in 2023, with results of those assessments expected to be released to the broader community working on conservation in the Lower Nicola Watershed in the near future.  

```{r rehab, echo = FALSE, results = 'asis'}
#| label: tbl-confirmed
#| tbl-cap: "Confirmed barriers"
#| warning: false
#| echo: false

data <- read.csv("data/confirmed_barriers.csv", check.names=FALSE)
ft <- flextable(data)
ft <- bg(ft, bg = "#008270", part = "header")
ft <- color(ft, color = "white", part = "header")
ft <- set_caption(ft)
ft |> autofit() 
```


```{r rehab, echo = FALSE, results = 'asis'}
#| label: tbl-datadeficient
#| tbl-cap: "Assessed structures that remain data deficient"
#| warning: false
#| echo: false

data <- read.csv("data/assessed_strucDD.csv", check.names=FALSE)
ft <- flextable(data)
ft <- bg(ft, bg = "#008270", part = "header")
ft <- color(ft, color = "white", part = "header")
ft <- set_caption(ft)
ft |> autofit() 

#ft <- format_flextable(ft)
#ft
```




```{r rehab, echo = FALSE, results = 'asis'}
#| label: tbl-rehabilitated
#| tbl-cap: "Rehabilitated Structures"
#| warning: false
#| echo: false

data <- read.csv("data/rehabilitated_barriers.csv", check.names=FALSE)
ft <- flextable(data)
ft <- bg(ft, bg = "#008270", part = "header")
ft <- color(ft, color = "white", part = "header")
ft <- set_caption(ft)
ft |> autofit() 

#ft <- format_flextable(ft)
#ft
```

## Glossary {-}

### Freshwater connectivity {-} 

The degree to which aquatic species can disperse and/or migrate freely through freshwater systems.  

#### Longitudinal connectivity {-} 

Connectivity of a stream along the upstream-downstream plane, including access to tributaries and spawning and rearing habitat. Longitudinal connectivity can be fragmented by physical barriers (e.g., anthropogenic, or natural structures) or by physiological limits of distribution for species (e.g., stream gradient, temperature, or flow requirements). 

#### Lateral connectivity {-} 

Connectivity of a stream bed to adjacent riparian wetlands, floodplains, side-channels, or off-channel features, including access to rearing and overwintering habitat. Lateral connectivity can be fragmented by physical barriers, channelization, armoring of the stream bed, or artificial flow regulation. 

#### Vertical connectivity {-}

Connectivity of a stream bed to groundwater/hyporheic zone, including access to oxygen-rich water and temperature refugia. Vertical connectivity can be fragmented by water withdrawals and anthropogenically induced changes to the hydrological, thermal, and sediment regimes of the watershed. 

#### Temporal connectivity {-}

Connectivity variability in any of the three spatial dimensions (longitudinal, lateral, vertical) based on changes in the natural flow regime over time. Variation in the temporal connectivity occurs naturally; however, fragmentation can be exacerbated through anthropogenically induced changes to the hydrological, thermal, and sediment regimes of the watershed. 


### Geographic Scope {-} 

The spatial bounds within which connectivity models are applied and fish-passage improvement efforts are focused. This area is refined after the accessibility model is developed to cover only streams that are naturally accessible to each focal species or guild (see naturally accessible streams below). 

### Focal Species  {-}

The ecologically and culturally important species (e.g., Chinook Salmon) for which habitat connectivity is being conserved and/or restored in the watershed. 

### Focal Guild  {-}

Two or more focal species grouped together by shared characteristics (e.g., migratory behavior, swimming ability, and habitat use).   

### Key Ecological Attribute (KEA)  {-}

An aspect of a focal species or guild’s biology or ecology that, if missing or altered, would lead to the loss of that focal species over time (e.g., proportion of connected spawning habitat). 

### Structure  {-}

A feature, either anthropogenic (e.g., culvert, dam) or a temporary natural feature (e.g., barrier beach, debris jam) that may fragment freshwater systems and limit the ability of species to disperse and/or migrate freely within their naturally accessible habitat.  Differentiated from more permanent natural features, such as waterfalls, that delineate naturally accessible waterbodies.  

### Set of structures  {-}

A set of two or more structures that must be addressed/considered together for restoration or rehabilitation when considering the total amount of connectivity to be gained during barrier prioritization. 


### Current Connectivity Status {-}

The measure of important habitat (i.e., amount, proportion, or index) that is currently connected (i.e., unaffected by structures with a status of unknown, partially passable, or impassable) for each focal species.  For non-diadromous species, indices such as the longest fragment or Dendridic Connectivity Index – potadromous (DCIp; Cote et al. 2009).  

### Connectivity Goal {-}

A SMART (Specific, Measurable, Achievable, Relevant, and Time-Bound) goal representing the desired future connectivity status of each focal species. 

### Structure Prioritization {-} 

An analysis that ranks structures (individually and/or in sets) by the amount of key habitat disconnected by each individual structure. Structures are ranked in descending order according connectivity gains (e.g., greatest amount of linear km of important habitat) if restored or rehabilitated. This analysis includes structures with a passability status of “unknown”, “partially passable”, or “impassable”.  
 
### Restoration {-}

Returning the ecological structure and function (e.g., natural channel-forming processes) of a site to pre-disturbance conditions. 

### Rehabilitation {-}

The partial recovery of ecosystem function. This is typical when improving passage by replacing one type of crossing with another that is passable but may still affect the streambed or channel-forming processes. 

### Rehabilitated barriers {-}

Structures for which connectivity for aquatic species was restored by altering or maintaining a structure that previously impeded movement. 

### Remediation {-}

A term for a form of restoration most often associated with mining, referring to the remediation of contaminated sediments. Often used as jargon in fish-passage discussions because it applies to both fish passage and barriers (i.e., both can be described as “remediated”).  Most fish-passage projects involve ecological rehabilitation, but it is inaccurate to write that a barrier was rehabilitated. Instead, fish passage should be described as rehabilitated, improved, or restored (if appropriate), and barriers described as removed or addressed. 

### Excluded structures {-}

Structures that were investigated during a project but were removed from consideration or the model. Reasons for their exclusion could be that they do not exist, are passable, or no key habitat was found upstream. 

### Assessed structures that remain data deficient {-}

Structures that have had some form of assessment but need further investigation before a decision can be made to categorize them as either confirmed barriers, excluded structures (passable, non-existent, no suitable upstream habitat), or rehabilitated structures.  



